Self-sustained carbon monoxide oxidation oscillations on size-selected platinum nanoparticles at atmospheric pressure

### 2012-04-30 paper.tex
Ready to go to Ib again

\section{Results and discussion}
An oscillation measurement is performed by mounting a sample in the setup
directly after cluster deposition and anodic bonding. Initially a light-off
CO-oxidation measurement is performed (shown in Figure~\ref{fgr:initial_treatment}) where
the temperature is increased at 3\,$^\circ$C/min until the sample ignites and
achieves full conversion of CO to CO$_2$. At a CO:O$_2$ ratio of $\sim0.08$ and a
geometrical Pt coverage of the reactor area of $\sim0.1\%$ light-off occurs at approximately
180\,$^\circ$C. The temperature ramp is typically continued up to
260\,$^\circ$C whereafter the temperature is decreased at 3\,$^\circ$C/min.
After the initial light-off the sample is cooled to room
temperature. When at room temperature the temperature is again increased at
3\,$^\circ$C/min until sustained oscillations occur. A typical initial
treatment is seen in Figure~\ref{fgr:initial_treatment}. Often the sample will
show oscillations either directly upon light-off or on the falling temperature
ramp.

@@

An example of a single very slow oscillation is seen in
Figure~\ref{fgr:full_oscillation}. After a short on-cycle (at $\sim$1030\,min) where
all CO is converted to CO$_2$ the reaction shows a fast deactivation. 
Notice the CO signal does not go to zero due to the cracking pattern of
CO$_2$ and CO background in the Quadrupole Mass Spectrometer (QMS).
After the deactivation the sample quickly regains
$\sim$30\% conversion (at $\sim$1040\,min). Hereafter the conversion increases
slowly over time until approximately 50\% conversion is reached (at
$\sim$1118\,min). Shortly after reaching 50\% conversion the sample again
ignites and converts all CO to CO$_2$. After a short full conversion period the
sample again deactivates and the cycle repeats itself.

@@

After prolonged measurements over several days an increase in oscillation
period was observed as shown in Figure~\ref{fgr:long_measurement}. Initially, the
oscillation period is $\sim$2.5 min and it increases linearly with time until 1500\,min
of total oscillation time. Hereafter the oscillations become more irregular
as shown in Figure~\ref{fgr:long_measurement} and in the lower panel of Figure~\ref{fgr:extracts}. At
experiment end the oscillation period is between 20\,min and 45\,min.

\begin{figure}
  \includegraphics[width=12cm]{summary_of_long_measurement.png}
  \caption{Summary of the oscillation period as a function of time for a sample
  oscillating under constant temperature, pressure and reactant composition.
  The sample went through a total of 439 oscillations in 4 days. For details on on
  the oscillations over this period please consult Figure~\ref{fgr:extracts}. Time is
  defined from experiment start and hence includes initial treatment. The inset
  shows the initial steady increase in oscillation period. The red lines indicate
  max and min oscillation time to guide the eye.}
  \label{fgr:long_measurement}
\end{figure}

Figure~\ref{fgr:extracts} shows extracts of mass spectrometry data of oscillations
extracted from the same four days measurement. Data from shortly after the
initial treatment, after 2800\,min of oscillation time and after 5900\,min are
shown. After 5900\,min the oscillations are more irregular which is consistent
with the data shown in Figure~\ref{fgr:long_measurement}. Also, oscillations in
between full conversion oscillations with smaller amplitude and much higher
frequency than the full on-off cycles become more visible as time progresses.
\begin{figure}
  \includegraphics[width=12cm]{extracts_from_very_long_oscillation.png}
  \caption{Extracts from a 4 days long experiment. The oscillation period
  becomes more irregular with time while the total integrated conversion
  remains constant. Furthermore, small oscillations in between full conversion
  cycles become more prominent as time progresses.} \label{fgr:extracts}
\end{figure}

The oscillation period increases with time and becomes increasingly less stable
as shown in Figure~\ref{fgr:extracts}. This phenomenon is consistent across all
measured samples. The large scatter in oscillation period can be attributed to
the smaller oscillations in between on/off cycles. Occasionally, these small
amplitude oscillations will trigger a full switch thus introducing more
short-period oscillations despite the trend of increasingly longer oscillation
period. A possible explanation of the phenomenon could be sintering of the
particles during the oxidation and reductions cycles. However, the ratio
between CO and CO$_2$ integrated over an full period remains constant
throughout the entire experiment time, i.e. the samples maintain a constant
average rate independent of oscillation frequency. Thus we do not expect
sintering to be the cause since an average loss in activity of the sample would
be expected.

@@

According to the proposed model in ref.~\citenum{Hendriksen2010} the bare metal
surface is less active than the oxidized surface. Oscillations originate from a
switch between an oxidized surface and the bare metal surface. Initially, with
CO present in the inlet gas a smooth metal surface is exposed due to the
general high adsorption energy of CO on metal surfaces. As the Pt nanoparticles
convert the CO to CO$_2$ the partial pressure of CO will decrease and a Pt
oxide will start to form in the high partial pressure of O$_2$ thus increasing
the rate. The model\cite{Hendriksen2010} suggests that the oxidized
nanoparticles will roughen during CO oxidation resulting in the formation of an
increasingly rough oxide surface. As the oxide becomes rougher the bare metal
surface is increasingly favored resulting in a return to the lower rate rough
metal state. As the rate decreases the CO partial pressure increases which
results in a transition from rough metal surface to a smooth metal surface.
When the sample is sufficiently smooth it will again oxidize and thus complete
the oscillation cycle.

@@

An interesting feature in the data is the very large temperature dependence of the
oscillation frequency as shown in Figure~\ref{fgr:temperature_dependence}. By
changing the temperature 5\,$^\circ$C the oscillation period was changed by more
than a factor of 10. The magnitude of the change is not completely consistent
across all measured samples but all samples show a very large temperature
dependence. In some cases a provoked change in temperature will
consistently turn on and off the oscillations. This is illustrated in the supplementary
material, Figure~S1.
This property agrees well an earlier suggested model \cite{Hendriksen2010} and
it is, to our knowledge, the first time the temperature dependence of
oscillations at atmospheric pressure has been measured.

@@

\subsection{CO concentration dependence}
The dependence of period of the oscillations on the the CO concentration was
investigated. The result shows that the period increases slightly with
increasing CO-concentration. However, the effect is small compared to the
general trend of slower oscillations as the experiment progresses. In
Figure~\ref{fgr:gas_dependence_summary} the oscillation periods are summarized and the
actual mass spectrometry data of the entire experiment are shown in Figure~\ref{fgr:gas_dependence}.

@@

\subsection{Proposed mechanism}
In this study Pt nanoparticles were investigated which, compared to a single
crystal or a thin film, have a very rough surface. It is to be expected that the
particles will initially favor the reduced state. As the reaction runs the gas
composition towards the outlet of the reactor will become more oxidizing
increasing the oxidation rate of the particles (illustrated in the bottom of
Figure~\ref{fgr:full_oscillation}). Gradually, as the particles towards the outlet
oxidizes the turn-over frequency will increase and the general CO concentration will
decrease promoting more and more particles to oxidize which will gradually be
seen in the QMS as an increase in CO$_2$ and a decrease in CO. As the CO
concentration decreases the light-off temperature will decrease hence
approaching the constant temperature of the sample resulting in a steep
increase in reactivity. It is important to note that the fraction of oxidized
particles needed to achieve full conversion is not known. We are currently
investigating new methods for gaining insight into the nature of the
nanoparticles in the reactor but the low coverage and the design of the reactor
makes this a challenging task. If all particles are not oxidized at light-off
they will now oxidize much faster due to the low CO-concentration. The further
increased rate will not be visible in the QMS since the reactor is already in
full conversion. As most of the particles are now oxidized they will become
more and more roughened and gradually return to the reduced state. Fewer and
fewer oxidized particles will be responsible for the activity increasing the
roughing rate of the remaining active particles. Just before the overall
activity drops only a fraction of the particles participates in the reaction
and the deactivation will happen very suddenly. Once these potentially few
particles eventually becomes sufficiently rough they will reduce and lose
activity hence completing the oscillation cycle. It should be noted that our
group has recently measured turnover frequencies of CO oxidation on Pt in
excess of $10^{6}$\,s$^{-1}$ per site corresponding to as little as $\sim$300 Pt
nanoparticles contributing to the reactivity just before the end of the high-
activity cycle. This would correspond to an area of 14\,ppm of the reactor
area suggesting that in situ studies are becoming a challenges as such a
minority of the reactor volume could account for all the reactivity. Attempts
to see any thermal signal from such small areas has so far been in vain.

\section{Conclusions}
Self-sustained CO oxidation oscillations at atmospheric pressure on Pt
nanoparticles ranging from from 3\,nm to 9\,nm in diameter have been measured with
high-quality mass-spectrometry. Full conversion was achieved on 0.1\,\%
geometrical Pt coverage and it is found likely that full conversion can be obtained on
extremely small amounts of Pt. It was found that the samples increased the
oscillation period over time and showed a strong temperature dependence of the
oscillation period. A model is proposed where the nanoparticles go through repetitive
oxidation/reduction cycles which change the change rate of CO oxidation on the
particles. The fact that extremely few NP may account for the activity make in
situ investigations a challenge.

>>>>

\section{Results and discussion}
An oscillation measurement is performed by mounting a sample in the setup
directly after cluster deposition and anodic bonding. Initially a light-off
CO-oxidation measurement is performed (shown in \ref{fgr:initial_treatment}) where
the temperature is increased at 3\,$^\circ$C/min until the sample ignites and
achieves full conversion of CO to CO$_2$. At a CO:O$_2$ ratio of $\sim0.08$ and a
geometrical Pt coverage of the reactor area of $\sim0.1\%$ light-off occurs at approximately
180\,$^\circ$C. The temperature ramp is typically continued up to
260\,$^\circ$C whereafter the temperature is decreased at 3\,$^\circ$C/min.
After the initial light-off the sample is cooled to room
temperature. When at room temperature the temperature is again increased at
3\,$^\circ$C/min until sustained oscillations occur. A typical initial
treatment is seen in \ref{fgr:initial_treatment}. Often the sample will
show oscillations either directly upon light-off or on the falling temperature
ramp.

@@

An example of a single very slow oscillation is seen in
\ref{fgr:full_oscillation}. After a short on-cycle (at $\sim$1030\,min) where
all CO is converted to CO$_2$ the reaction shows a fast deactivation. 
Notice the CO signal does not go to zero due to the cracking pattern of
CO$_2$ and CO background in the Quadrupole Mass Spectrometer (QMS).
After the deactivation the sample quickly regains
$\sim$30\% conversion (at $\sim$1040\,min). Hereafter the conversion increases
slowly over time until approximately 50\% conversion is reached (at
$\sim$1118\,min). Shortly after reaching 50\% conversion the sample again
ignites and converts all CO to CO$_2$. After a short full conversion period the
sample again deactivates and the cycle repeats itself.

@@

After prolonged measurements over several days an increase in oscillation
period was observed as shown in \ref{fgr:long_measurement}. Initially, the
oscillation period is $\sim$2.5 min and it increases linearly with time until 1500\,min
of total oscillation time. Hereafter the oscillations become more irregular
as shown in \ref{fgr:long_measurement} and in the lower panel of \ref{fgr:extracts}. At
experiment end the oscillation period is between 20\,min and 45\,min.

\begin{figure}
  \includegraphics[width=12cm]{summary_of_long_measurement.png}
  \caption{Summary of the oscillation period as a function of time for a sample
  oscillating under constant temperature, pressure and reactant composition.
  The sample went through a total of 439 oscillations in 4 days. For details on on
  the oscillations over this period please consult \ref{fgr:extracts}. Time is
  defined from experiment start and hence includes initial treatment. The inset
  shows the initial steady increase in oscillation period. The red lines indicate
  max and min oscillation time to guide the eye.}
  \label{fgr:long_measurement}
\end{figure}

Figure~\ref{fgr:extracts} shows extracts of mass spectrometry data of oscillations
extracted from the same four days measurement. Data from shortly after the
initial treatment, after 2800\,min of oscillation time and after 5900\,min are
shown. After 5900\,min the oscillations are more irregular which is consistent
with the data shown in \ref{fgr:long_measurement}. Also, oscillations in
between full conversion oscillations with smaller amplitude and much higher
frequency than the full on-off cycles become more visible as time progresses.
\begin{figure}
  \includegraphics[width=12cm]{extracts_from_very_long_oscillation.png}
  \caption{Extracts from a 4 days long experiment. The oscillation period
  becomes more irregular with time while the total integrated conversion
  remains constant. Furthermore, small oscillations in between full conversion
  cycles become more prominent as time progresses.} \label{fgr:extracts}
\end{figure}

The oscillation period increases with time and becomes increasingly less stable
as shown in \ref{fgr:extracts}. This phenomenon is consistent across all
measured samples. The large scatter in oscillation period can be attributed to
the smaller oscillations in between on/off cycles. Occasionally, these small
amplitude oscillations will trigger a full switch thus introducing more
short-period oscillations despite the trend of increasingly longer oscillation
period. A possible explanation of the phenomenon could be sintering of the
particles during the oxidation and reductions cycles. However, the ratio
between CO and CO$_2$ integrated over an full period remains constant
throughout the entire experiment time (supplementary, Figure~S2), i.e. the samples maintain a constant
average rate independent of oscillation frequency. Thus we do not expect
sintering to be the cause since an average loss in activity of the sample would
be expected.

@@

According to the proposed model in ref.~\citenum{Hendriksen2010} the bare metal
surface is less active than the oxidized surface. Oscillations originate from a
switch between an oxidized surface and the bare metal surface. Initially, with
CO present in the inlet gas a smooth metal surface is exposed due to the
general high adsorption energy of CO on metal surfaces. As the Pt nanoparticles
convert the CO to CO$_2$ the partial pressure of CO will decrease and a Pt
oxide will start to form in the high partial pressure of O$_2$ thus increasing
the rate. The model suggests that the oxidized
nanoparticles will roughen during CO oxidation resulting in the formation of an
increasingly rough oxide surface. As the oxide becomes rougher the bare metal
surface is increasingly favored resulting in a return to the lower rate rough
metal state. Since the surface mobility is higher for the reduced surface\cite{Hendriksen2010},
the roughness of the surface will gradually decrease and result
in a transition from rough metal surface to a smooth metal surface.
When the sample is sufficiently smooth it will again oxidize and thus complete
the oscillation cycle.

@@

An interesting feature in the data is the very large temperature dependence of the
oscillation frequency as shown in \ref{fgr:temperature_dependence}. By
changing the temperature 5\,$^\circ$C the oscillation period was changed by more
than a factor of 10. The magnitude of the change is not completely consistent
across all measured samples but all samples show a very large temperature
dependence. In some cases a provoked change in temperature will
consistently turn on and off the oscillations. This is illustrated in the supplementary
material, Figure~S1.
This property agrees well an earlier suggested model \cite{Hendriksen2010} and
it is, to our knowledge, the first time the temperature dependence of
oscillations at atmospheric pressure has been measured.

@@

\subsection{CO concentration dependence}
The dependence of period of the oscillations on the the CO concentration was
investigated. The result shows that the period increases slightly with
increasing CO-concentration. However, the effect is small compared to the
general trend of slower oscillations as the experiment progresses. In
\ref{fgr:gas_dependence_summary} the oscillation periods are summarized and the
actual mass spectrometry data of the entire experiment are shown in \ref{fgr:gas_dependence}.

@@

\subsection{Proposed mechanism}
In this study Pt nanoparticles were investigated which, compared to a single
crystal or a thin film, have a very rough surface. It is to be expected that the
particles will initially favor the reduced state. As the reaction runs the gas
composition towards the outlet of the reactor will become more oxidizing
increasing the oxidation rate of the particles (illustrated in the bottom of
\ref{fgr:full_oscillation}). Gradually, as the particles towards the outlet
oxidizes the turn-over frequency will increase and the general CO concentration will
decrease promoting more and more particles to oxidize which will gradually be
seen in the QMS as an increase in CO$_2$ and a decrease in CO. As the CO
concentration decreases the light-off temperature will decrease hence
approaching the constant temperature of the sample resulting in a steep
increase in reactivity. It is important to note that the fraction of oxidized
particles needed to achieve full conversion is not known. We are currently
investigating new methods for gaining insight into the nature of the
nanoparticles in the reactor but the low coverage and the design of the reactor
makes this a challenging task. If all particles are not oxidized at light-off
they will now oxidize much faster due to the low CO-concentration. The further
increased rate will not be visible in the QMS since the reactor is already in
full conversion. As most of the particles are now oxidized they will become
more and more roughened and gradually return to the reduced state. Fewer and
fewer oxidized particles will be responsible for the activity increasing the
roughing rate of the remaining active particles. Just before the overall
activity drops only a fraction of the particles participates in the reaction
and the deactivation will happen very suddenly. Once these potentially few
particles eventually becomes sufficiently rough they will reduce and lose
activity hence completing the oscillation cycle. It should be noted that our
group has recently measured turnover frequencies of CO oxidation on Pt in
excess of $10^{6}$\,s$^{-1}$ per site corresponding to as little as $\sim$300 Pt
nanoparticles contributing to the reactivity just before the end of the high-
activity cycle. This would correspond to an area of 14\,ppm of the reactor
area suggesting that in situ studies are becoming a challenges as such a
minority of the reactor volume could account for all the reactivity. Attempts
to see any thermal signal from such small areas has so far been in vain.

\section{Conclusions}
Self-sustained oscillations have been found to be of great interest in nature
and here we demonstrate a systems that shows how nanoparticles can be brought
into this regime with high reliability forming a basis for further deep
understanding of this phenomena.

Self-sustained CO oxidation oscillations at atmospheric pressure on Pt
nanoparticles ranging from from 3\,nm to 9\,nm in diameter have been measured with
high-quality mass-spectrometry. Full conversion was achieved on 0.1\,\%
geometrical Pt coverage and it is found likely that full conversion can be obtained on
extremely small amounts of Pt. It was found that the samples increased the
oscillation period over time and showed a strong temperature dependence of the
oscillation period. A model is proposed where the nanoparticles go through repetitive
oxidation/reduction cycles which change the change rate of CO oxidation on the
particles. The fact that extremely few NP may account for the activity make in
situ investigations a challenge.
